### 2018-04-19 papers/synth/abstract.tex
asd

Vesicle Traffic Systems(VTS) are the material transport mechanisms
among the compartments inside the biological cells.
%
The compartments are viewed as nodes that are labeled with the containing
chemicals and the transport channels are labeled edges between the
nodes.
%
Understanding VTSs is an ongoing area of research and for many cells
they are partially known.
%
For examples, there may be undiscovered edges, nodes, or their labels
in a VTS of a cell.
%
It has been speculated that there are properties that the VTSs must satisfy.
%
For example stability, i.e., every chemical that is leaving a compartment comes back.
%
Many synthesis questions may arise in this scenario, where we want
to complete a partially known VTS under a given property.
%
In the paper, we present novel encodings of the above questions
into the QBF satisfiability problems.
%
We have implemented the encodings in a highly configurable tool
and  apply to a couple of found-in-nature VTSs and 
several synthetic graphs.
%
Our results demonstrates that our method can scale upto the
graphs of interest.
%

>>>>

Vesicle Traffic Systems(VTSs) are the material transport mechanisms
among the compartments inside the biological cells.
%
The compartments are viewed as nodes that are labeled with the
containing chemicals and the transport channels are similarly labeled
edges between the nodes.
%
Understanding VTSs is an ongoing area of research and for many cells
they are partially known.
%
For example, there may be undiscovered edges, nodes, or their labels
in a VTS of a cell.
%
It has been speculated that there are properties that the VTSs must
satisfy.
%
For example, stability, i.e., every chemical that is leaving a
compartment comes back.
%
Many synthesis questions may arise in this scenario, where we want
to complete a partially known VTS under a given property.
%
In the paper, we present novel encodings of the above questions
into the QBF(quantified Boolean formula) satisfiability problems.
%
We have implemented the encodings in a highly configurable tool
and  apply to a couple of found-in-nature VTSs and 
several synthetic graphs.
%
Our results demonstrates that our method can scale upto the
graphs of interest.
%

### 2018-04-19 papers/synth/bio.tex
asd

In this section, we will formally define VTSs as a labeled graph
and the discussed conditions on the graphs.
%
Subsequently, we will present the SMT encoding of the search problems
for the graphs that satisfy the conditions.

\subsection{Model}

We model VTSs as labeled directed graphs.
The graph labels both nodes and edges with sets of molecules to denote
the set of molecules present in them.
The graph is formally defined as follows.

@@

We will also consider several variations of the model.
%
For example, unique edge between two nodes, activity of molecules is
not constrained by $\nodef$ and $\edgef$, etc.
%

@@

\paragraph{Search problem}
Based upon earlier discussion, we need to answer the following search
question among VTSs.
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols| = \mu$, and
$G$ is not $k$-connected.

\subsection{Boolean satisfiability of the search problem}

We translate the search problem into a Boolean satisfiability
problem and use SMT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterwards, we will present the constraints corresponding to the
properties.

\subsubsection{Variables for VTS description}
%
We assume that size of the graph is $\nu$ and number of molecules is
$\mu$.
%
Furthermore, we also limit the maximum number $\pi$ of edges present
between two nodes.
%
Here, we list the vector of Boolean variables and uninterpreted function symbols
that encode the VTSs.
\begin{enumerate}

\item Boolean variable $n_{i,m}$ indicates if $m \in \nlabel(i)$
\item Boolean variable $e_{i,j,q}$ indicates if $q$th edge exists between $i$ and $j$.
\item Boolean variable $e_{i,j,q,m}$ indicates if $q$th edge between $i$ and $j$ contains $m$.
\item Boolean variable $p_{m,m'}$ indicates if $(m,m') \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}
We also have auxiliary Boolean variables that will help us encode the well-fused and stability properties 
\begin{enumerate}
\item $a_{i,m}$ indicates that molecule $m$ is active at node $i$, i.e., $\nodef(m,L(i))$
  holds
\item $b_{i,j,q,m}$ indicates that molecule $m$ is active at $q$th edge $(i,M',j)$ between $i$ and $j$, i.e., $\edgef(m,M')$ holds
\item $r_{i,j,m,p}$ indicates if there is an $m$-path from
  $i$ to $j$ of length less than or equal to $p$
\end{enumerate}
For $k$-connected property, we also use the following auxiliary Boolean variables
\begin{enumerate}
\item $d_{i,j,q}$ indicates $q$th edge between $i$ and $j$ is dropped
\item $r'_{i,j}$ indicates if there is a path from $i$ to $j$ in the modified VTS
\end{enumerate}

We will describe the Boolean constraints that encode VTSs in several categories.
%
In the end, we will present in a table the constraints needed for the
model variants.
%
To avoid cumbersome notation, we will not explicitly write the ranges of the indexing
in the constraints.
%
$i$ and $j$ will range over nodes.
%
$m$ will range over molecules.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints~\eqref{eq:f0} and~\eqref{eq:c3}
to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align}
  \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}\tag{V1}\label{eq:f0}
\end{align}
If a molecule is active on an edge, it should be present on the edge.
%
\begin{align}
  \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\tag{V2}\label{eq:f1}
\end{align}
A molecule should be present to be active on a node.  
\begin{align}
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  \tag{V3}\label{eq:c4}
\end{align}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )\tag{V4}\label{eq:c0}
\end{align}
Self edges are not allowed. 
\begin{align}
   \bigwedge\limits_{i,q} \neg e_{i,i,q}\tag{V5}\label{eq:c2}
\end{align}
We fix first half as Q-SNAREs and rest as R-SNAREs and set diagonal blocks in pairing matrix to be 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
\begin{align}
  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
  \tag{V6}\label{eq:c3}
\end{align}

\subsubsection{Well-fused constraints}
Constraint~\eqref{eq:fuse1} encodes that each edge must fuse with
its destination node.
%
Constraint~\eqref{eq:fuse2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align}
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies \bigvee_{m,m^{\prime}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \tag{V7}\label{eq:fuse1}  \\
\bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies \neg \bigvee_{j \neq j^{\prime}, m^{\prime\prime}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \tag{V8}\label{eq:fuse2}  
\end{align}

% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

% F4: For an edge to be valid, at least one SNARE pair on the vesicle and target compartment must be active and have a non-zero entry in the pairing matrix.

% F5: To ensure that fusion respects the graph structure by the edge under consideration, it should not be possible to fuse with any other node.

\subsubsection{Regulation on nodes and edges}
We are considering several models that differ in constraints on
the activity of molecules.
%
We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
the varying constraints.
%
The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} = a_{i,m}    \tag{Ann}\label{eq:ann}
\end{align}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\tag{Anb}\label{eq:anb}
\end{align}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\tag{Aen}\label{eq:aen}
\end{align}
The following constraint encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align}
   \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
  \tag{Aeb}\label{eq:aeb}
\end{align}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align}
   \bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \limplies  \bigwedge_{m' \neq m} (p_{m,m'} \limplies e_{i,j,q,m'})] \lequiv (\neg b_{i,j,q,m} \land\;\;\;\;  \bigwedge_{\mathclap{m' \neq m, p_{m,m'}}} \neg b_{i,j,q,m'})
  \tag{Aep}\label{eq:aep}
\end{align}
%
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align}
  \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
  \tag{R1}\label{eq:reach1}
\end{align}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
  \tag{R2}\label{eq:reach2}
\end{align}

% Boolean variable
% F3 and F2 are used to model these constraints.

% % \textbf {Reachability definition and stability condition.}
% We have encoded stable condition by using reachability definition. \newline

% F3: Stability condition. For every leaving molecule, the source node is reachable from the target node with that molecule present, in maximum p steps.

% F2: Reachability definition. Node j is reachable from node i with kth molecule present in maximum p steps if either there is a direct edge between i-j with that molecule present or there is a direct edge between $i^{\prime \prime}$ $(i \neq i^{\prime \prime})$ and j with kth molecule present on that edge and $i^{\prime \prime}$ are reachable from i in p steps.

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Constraints~\ref{eq:drop1}-\ref{eq:drop4}
encode the relevant constraints for reachability
in the modified VTS. 
%
The following constraints encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following constraint, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we search for disconnected modified VTS,
the following constraint encodes that there are 
unreachable pair of nodes in the underlying undirected graph.
% nodes such that
% there is no path between them .
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

The key improvement in our work over the earlier tool is the
encoding of reachability, which was done using enumeration of paths.
%
In the current work, we have encoded reachability in two different
ways in constraints~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined in~\eqref{eq:drop3} and has
trivial solutions by making all $r'$s true.
%
However, the trivial solutions are disallowed by constraint~\eqref{eq:drop4} and we find
only the solutions that captures the evidence of unreachability.
%
On the other hand,
we have added length of paths in our reachability encoding in constraint~\eqref{eq:reach1},
which needs relatively more auxiliary variables.
%
This is because constraint~\eqref{eq:reach2} has only positive
occurrences of the reachability variables and if we had defined
$r$s in ~\eqref{eq:reach1} without paths,
the circular dependencies in the recursive definitions of $r$s
may have resulted in spurious satisfying assignments that
do not encode reachability.
%
By adding the path length, we break the circular dependencies, the
constraint remains polynomial in size, and satisfying assignments only
correspond to correct reachability.

% We can go down or up using -C \_ option.  \newline

\subsection{Complete formula}
The complete QBF looks like the following.
\begin{align}
\exists \, e_{i,j,q} \forall \,d_{i,j,q}, r^{\prime}_{i,j,q} \, (D1 \land D2 \land D3 \land D4)  \, \land \notag\\  \forall\, n_{i,k}, a_{i,k}, e_{i,j,k,q}, b_{i,j,k,q}, r_{i,j,k}, p_{k,k^{\prime}} \, \neg (V1 \land V2 \land ... V8 \land A_{n} \land A_{e})  \label{eq:final}
\end{align}

\subsection{SMT problems for different variants}
%\ankit{Bring model variant table here}

We have encoded the following variants of VTSs.
%
The variants are due to different combinations of constraints on the
activity of molecules on the nodes and edges.
%
\begin{enumerate}[label=\Alph*]
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
\end{enumerate}
%
% For the version A, every present molecule is in active state.
%
In the table~\ref{tab:grph}, we present the constraints involved in each version.
%
The column two of the table shows that the constraints V1-V8, R1-R2, and D1-D4 are present
for every variant.
%
The column three of the table lists the constraints that are different among the variants.
%
%
One of the restriction is where the activity of the present molecule is dependent on the presence of the other molecules.
%
For example in version B, D, F activity of the molecules on the node is a boolean function of the presence of other molecules on that node; Anb.
%
Similarly for the case of the edge in version C and D; Aeb.
%
In the case of versions F the activity of the molecules on the edges
is described by pairing matrix; Aep.

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SMT solver to find
VTSs that belong to the variants.

% We have applied the tool for the above variations for searching the
% VTSs that satisfies their corresponding properties. 
%
% After creating the constraint using different Boolean variables 
% we ship them directly to the solver.
% %
% In case of a satisfying (SAT) assignment of the added constraints the
% Z3 theorem prover gives a model with the variable assignment.
% %
% In case no possible satisfying assignment the solver returns UNSAT.

%
%

% \subsection{Creating a solver and solving these constraints}
% We will create a solver and add these build constraint using these constraints to that and then check whether there a satisfiable assignment that satisfies all these constraints.

% \begin{figure}[ht]
% %\vspace{-3mm}
% \begin{verbatim}
% # Create a solver
% s = Solver()
% # Add these constraints to it.
% s.add(A0, A1, C0, C1, C2, C4, C5, F0, F1, F2, F3, F4, F5, D0, D1, D2, D3)

% # Check for the satisfiability assignment.
% print s.check()   
% \end{verbatim}
% \caption{Solving for these constraints.}
% \label{code:motivate}
% \end{figure}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

>>>>

In this section, we will present the model of VTS from~\cite{PloS,SASB}.
%
We will also present the constrains and properties on the VTSs.   
%
We model a VTS as labelled graph along with assisting pairing matrices and
activating functions.

@@

% We will also consider several variations of the model.
% %
% For example, unique edge between two nodes, activity of molecules is
% not constrained by $\nodef$ and $\edgef$, etc.
%

@@

### 2018-04-19 papers/synth/intro.tex
asd

Vesicle Traffic Systems are the transport mechanisms inside the cells.
%
Different compartments are viewed as nodes and transport channels are
nodes between the graphs.

\ashu{Speculative}
Subsequently, it can be potentially useful in guiding biologists

>>>>

Vesicle Traffic Systems(VTS) are the material transport mechanisms
among the compartments inside the biological cells~\cite{vtsIntro}.
%
\ashu{Describe VTS in detail}
%
Vesicle Traffic Systems are the transport mechanisms inside the cells.
%
Different compartments are viewed as nodes and transport channels are
nodes between the graphs.

\ashu{Make a case for incomplete knowledge of VTSs.}
%Now we will work on
Since VTSs are partially known, 
%
the labels may not have been fully observed.

\ashu{Make a case for expected properties!!}
Furthermore, it is a matter of debate what properties the VTSs should
have, such as stability, i.e., every chemical that is leaving a
compartment comes back.

\ashu{Speculative: Make a case for completing partial parts of VTSs!!}

We model a VTS as a labelled graph.
%
Each compartment of VTS is a node of the graph and and the
transport channels between the nodes is an edge.
%
The set of chemicals present in a compartment is the label of of
the compartment.
%
Similarly, the set of transported chemicals carried by an edge is the label
on the edge.
%
%

There are several constraints that limits

There are several constraints that encode the interaction.

The labels on the nodes
